---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

This dataset is an account of all of the checkouts for stories from the author 
Sarah J. Maas. Author Maas is known for many great romance and fantasy novels, 
so I am a fan of her works and wanted to look at the data over the past couple 
years of just her books. During Covid-19 there was a rise of this "Booktok" 
hashtag on the app Tiktok, many videos of people suggesting books for others to 
read. I have come across this hashtag multiple times during quarantine, most of 
the videos being romance or fantasy recommendations. I know that one series 
from Sarah J. Maas was heavily recommended on Tiktok, so I am looking at the 
data with questions surrounding the influence of Tiktok on book checkout numbers,
and overall how quarantine affected fantasy and romance checkout numbers. Why 
am I curious about this in particular: many have used apps like Tiktok and read 
novels of fantasy, romance, or distopian genres as a means of escape during 
Covid qurantine times, I simply want to see if this is true for books from an 
author like Sarah J. Maas!

### Summary Information


According to a couple of articles from multiple news sources (like the Guardian), 
Booktok, a hashtag page on the app Tiktok started and became trending after 
August 2020 when a person's book recommendation Tiktok went viral. It has become 
a page of popular book recommendations (mostly romance) for people to find out 
about books. Sarah J. Maas's books like A Court of Thorns and Roses have excelled 
in the romance category in terms of recommendations on Tiktok.

Looking at the data, we have to take into account the rise of eBooks from 
intense quaranting during the Covid pandemic (libraries and bookstores closed), 
the increased number of books read in general from people being stuck in quarantine, 
as well as the rise in demand for romance and fantasy novels (from 
recommendations on Tiktok). The average number of checkouts per item in the 
dataset is _17.6_ and looking at the number of checkouts per year we 
got:
2012: __*80*__ 2013: __*291*__ 2014: __*752*__ 2015: __*1553*__ 2016: __*2563*__
2017: __*3405*__ 2018: __*4431*__ 2019: __*5595*__ 2020: __*6238*__ 
2021: __*11212*__ 2022: __*18370*__ 2023: __*2141*__ 
(We are not looking at 2023 too much since the year is not over yet)

Looking at this data we can definitely see a steady increase each year, and a 
sharp increase (almost doubles from 2020 to 2021) which can imply the theories 
regarding Covid-19 and quarantine, as well as Tiktok. Looking more closely at 
2020 after the month of August (when Booktok really kicked off), there is a steady 
increase in the numbers after those months. The month with the maximum eBook 
checkouts in 2020 was _December_, further proving the previous statement of the 
increase. Looking at the differences between ebooks and paper books, we find 
that the number of checkouts for regular books each year (from 2019 to 2021) goes 
from _1192 to 366 to 958_ (differences calculated as: _-826, 592_). 
This suggests that the pandemic and quarantining had a significant effect on the 
number of checked out physical books (probably due to libraries and bookstores 
being closed, and people not wanting to or allowed to leave their homes). 
As for eBooks, the numbers (from 2019 to 2021) go from _4403 to 5872 to 10254_ 
(differences calculated as: _1469, 4382_). There is a definite increase between 
these years, suggesting that quarantine increased eBook purchases and pushed people 
to try out eBooks for the first time (and ending up using that method for traveling 
etc so forth due to its comfortability). It's hard to say whether or not Tiktok 
really impacted these numbers, considering the popularity of some of Sarah's 
novels, but one thing for sure we can tell from these calculations is that Covid-19 
had a huge impact on the number of books people read as well as the use of eBooks.

Questions used:
- What is the average number of checkouts for each item?
- What is the number of checkouts each year?
- What is the number of eBook vs regular book checkouts in the year 2019?
- "" 2020
- "" 2021
- How did the numbers change between these years? (differences)
- What is the month with the most checkouts for ebooks in 2020?
- What are the checkout numbers in 2020 by month?


### The Dataset

This dataset focuses on checkouts of any Sarah J. Maas books from the Public 
Library in Seattle each month from the years 2012-2023. This dataset refreshes 
every month, published and collected by the Seattle Public Library (SPL). They 
take summary data from their catelogs on checked out books and history of those 
numbers. We have to take into account accessibility mainly, as some of her books
were very popular and hard to get a hold of from libraries. This could lessen 
the number of checked out physical books if there were not many copies available,
or if there was a long line to hold a book. This data set does not account for 
those that downloaded pdfs of the books off of a website, or those who borrowed 
the book from someone who already has it (example: borrowed book from sister's 
collection to read). The dataset also does not precisely show the correlation 
with the rise of booktok on tiktok for book recommendations, but can give us an 
idea of how covid-19 quarantining contributed to the numbers. This data collection 
started April 2005, however we are only looking at Sarah J. Maas books, which 
starts during 2012. This data was probably collected to look at trends of eBooks 
versus physical books of different genres. With the rise of technology and its 
uses, it raises questions about the future of libraries, and this data helps 
visualize this transformation over the past decade.



### Checkout Trends Over Time Chart
This chart shows the number of checkouts of Sarah J. Maas books over the years.
This chart was included to show the correlation between the pandemic and social
media apps with checkout trends. From this chart we can conclude that there is
indeed a sharp increase in numbers after 2020 (when quarantine really started
and #Booktok kicked off) and a decline in numbers after quarantine ended. This 
suggests that fantasy and romance novels were quite in demand during quarantine
as a possible means of escape or way to spend time stuck inside.


``````{r, echo = FALSE, code = readLines("chart1_example.R")}
```

### Ebook vs Physical Book Trends Over Time Chart
This chart shows the number of checkouts for Ebooks vs Books over time. This chart
was included to show the correlation between the pandemic and the increase of 
use of Ebooks. With the rise of technology and its uses its hard to precisely tell
if the pandemic alone contributed to this increase, but it is still interesting to
see in the chart below. This chart suggests that the number of physical books
checked out stayed roughly the same, increasing slightly during quarantine (makes
sense considering the amount of time people spent indoors), but ebooks mostly have
shined during and throughout this whole pandemic with the sharp increase following 2020.
Include a chart. Make sure to describe why you included the chart, and what patterns emerged


``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Bar Plot
This Bar plot shows the different series/titles that Sarah J. Maas has written 
and the number of checkouts for these novels. The TOG (throne of glass) is a
special side short stories series that I included separately. What is surprising
looking at this graph is that the Court Series (ACOTAR) has fewer books than the
Throne of Glass Series, yet has much more checkouts. This can support the 
theories of this report surrounding the increased demand of romance and fantasy 
books during Covid (ACOTAR is a fantasy romance series). This also further 
contributes to the theory surrounding Tiktok's Booktok page influence, as ACOTAR 
is a popular series with its own hashtag on the app. 

``````{r, echo = FALSE, code = readLines("barplot.R")}
```


